---
title: "Hardest Professors at MSU"
output: html_notebook
---



```{r}
library(tidyverse)
glimpse(tidy_grades)
```

```{r}
tidy_grades %>%
  group_by(instructor, subject_code) %>%
  filter(sum(quantity)>100,
         course_code >= 200 & course_code <= 800) %>%
  summarise(mean_grade = sum(grade*quantity)/sum(quantity),
            median_grade = median(rep(grade, quantity))) %>%
  arrange(mean_grade)
```

```{r}
profs_mean_grade_desc <- tidy_grades %>%
  group_by(instructor, subject_code, course_code) %>%
  filter(sum(quantity)>100,
         course_code >= 300 & course_code <= 800) %>%
  summarise(mean_grade = sum(grade*quantity)/sum(quantity),
            median_grade = median(rep(grade, quantity))) %>%
  arrange(mean_grade)

profs_mean_grade_desc
```
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.


```{r}
tidy_grades %>%
  filter(subject_code == "FI",
         course_code == 457) %>%
  summarise(mean_grade = sum(grade*quantity)/sum(quantity),
            median_grade = median(rep(grade, quantity))) %>%
  glimpse()
```


EC 406 is "Economic Analysis of Russia" taught by Professor Susan J. Linz.
```{r ec 406 distribution}
tidy_grades %>%
  filter(subject_code == "EC" & course_code == "406") %>%
  group_by(grade) %>%
  summarise(freq = sum(quantity)) %>%
  ggplot(aes(grade, freq)) +
  geom_histogram(stat = "identity")
```

* Susan Linz (EC)
* Jeff Biddle (EC)
* Randall Schaetzl (GEO)
* Christopher J Paul (ME)
* Helen Mayer (VM)
* Patrick M Muzzall (ZOL)
* John D Molloy (ISS)
* Ralph Fogwell (ANS)
* Jon Pumplin (PHY)
* Corwin D Smidt (PLS)
* Yimin Xiao (STT)
* Brendan P Mullan (SOC)
