---
title: "Hardest Professors at MSU"
output: html_notebook
---



```{r}
library(tidyverse)
glimpse(tidy_grades)
```

```{r}
tidy_grades %>%
  group_by(instructor, subject_code) %>%
  filter(sum(quantity)>100,
         course_code >= 200 & course_code <= 800) %>%
  summarise(mean_grade = sum(grade*quantity)/sum(quantity),
            median_grade = median(rep(grade, quantity))) %>%
  arrange(mean_grade)
```

```{r}
tidy_grades %>%
  group_by(instructor, subject_code, course_code) %>%
  filter(sum(quantity)>100,
         course_code >= 300 & course_code <= 800) %>%
  summarise(mean_grade = sum(grade*quantity)/sum(quantity),
            median_grade = median(rep(grade, quantity))) %>%
  arrange(mean_grade)
```
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.
```{r}
#biddle
tidy_grades %>% 
  filter(instructor == "JEFF E BIDDLE") %>%
  group_by(instructor, subject_code) %>%
  summarise(mean_grade = sum(grade*quantity)/sum(quantity),
            median_grade = median(rep(grade, quantity))) 
  glimpse()
```

```{r}
tidy_grades %>%
  filter(subject_code == "FI",
         course_code == 457) %>%
  summarise(mean_grade = sum(grade*quantity)/sum(quantity),
            median_grade = median(rep(grade, quantity))) %>%
  glimpse()
```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

* Susan Linz (EC)
* Jeff Biddle (EC)
* Randall Schaetzl (GEO)
* Christopher J Paul (ME)
* Helen Mayer (VM)
* Patrick M Muzzall (ZOL)
* John D Molloy (ISS)
* Ralph Fogwell (ANS)
* Jon Pumplin (PHY)
* Corwin D Smidt (PLS)
* Yimin Xiao (STT)
* Brendan P Mullan (SOC)
